Title

### 2018-12-19 MAESTROeX_I/paper.tex
Include computation of etarho and psi in numerical scheme

@@

\item[Step 0.] {\em Initialization}
The initialization step only occurs at the beginning of the simulation.
The initialization step computes initial values for $\pi^{-\myhalf}$, $S^0$, and
$S^1$ for use in time stepping routine.  We pass through {\bf Step 1} -- {\bf Step 12},
but only keep these values.
The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0,
\rho_0^0, p_0^0$, and $\overline{\Gamma}_1^0$ are specified from the problem-dependent
initial conditions.  The initial time step, $\dt^0$, is computed as in
Paper III.

@@

Using the equation below,
%
\begin{equation}
\frac{\partial\Ub}{\partial t} = -\Ub\cdot\nabla\Ub - \frac{1}{\rho}\nabla\pi - \frac{\rho-\rho_0}{\rho}g\eb_r, \label{eq:u_update}
\end{equation}
%
we compute time-centered edge velocities, $\uadvonedag$, using
$\Ub^n$.  The $\dagger$ superscript refers to the
fact that the predicted velocity field does not satisfy the divergence
constraint.  We then construct $\uadvone$ from $\uadvonedag$
using a MAC projection.
We note that $\uadvone$ satisfies the divergence constraint
\begin{equation}
\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \frac{1}{\overline{\Gamma}_1^np_0^n}\frac{p_0^n-p_0^{n-1}}{\Delta t^{n-1}} + \chi^n\right),
\end{equation}
\begin{equation}
 \beta_0^n = \beta_0 \left(\rho_0^n, p_0^n, \overline{\Gamma}_1^n\right),
\end{equation}
\begin{equation}
\chi^n = \frac{f}{\overline{\Gamma}_1^n p_0^n}\frac{p_{\rm eos}^n(\rho,h,X_k) - p_0^n}{\Delta t^n}
\end{equation}
where $\beta_0$ is computed as described in Appendix C of Paper III.
\MarginPar{add volume discrepancy stuff}

@@

\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$.
\begin{description}
\item For planar geometry,
\begin{eqnarray}
(\rho h)^{(2),\star}
&=& (\rho h)^{(1)} \nonumber \\
&&- \dt^n \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \dt^n \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} \nonumber \\
&& + \dt^n \left( \frac{p_0^n - p_0^{n-1}}{\dt^{n-1}} \right),
\end{eqnarray}

\item For spherical geometry,
\begin{eqnarray}
(\rho h)^{(2),\star}
&=& (\rho h)^{(1)} \nonumber \\
&&- \dt^n \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \dt^n \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) - p_0^{n} \nabla \cdot \uadvone \right \} \nonumber \\
&& + \dt^n \left( \frac{p_0^n - p_0^{n-1}}{\dt^{n-1}} \right),
\end{eqnarray}
\end{description}

@@

Note that during the first time step we use $\dt^n$ instead of $\dt^{n-1}$.
Then, for each Cartesian cell where $\rho^{(2),\star} < \rho_\mathrm{cutoff}$,
we recompute enthalpy using
\begin{equation}
(\rho h)^{(2),\star} = \rho^{(2),\star}h\left(\rho^{(2),\star},p_0^{n+1,\star},X_k^{(2),\star}\right).
\end{equation}

@@

The procedure to construct $\uadvtwodag$ is identical to the procedure
for computing $\uadvonedag$ in {\bf Step 3}, but uses
the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
We note that $\uadvtwo$ satisfies the divergence constraint
\begin{equation}
\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph} \left(S^{\nph} - \frac{1}{\overline{\Gamma}_1^{n+\myhalf}p_0^n}\frac{p_0^{n+1,\star}-p_0^n}{\Delta t^n} + \chi^{n+\myhalf}\right),
\end{equation}
\begin{equation}
\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2};
\qquad
 \beta_0^{n+1,\star} = \beta_0 \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \overline{\Gamma}_1^{n+1,\star}\right),
\end{equation}
\begin{equation}
\overline{\Gamma}_1^{\nph} = \frac{ \overline{\Gamma}_1^n +  \overline{\Gamma}_1^{n+1,\star} }{2}.
\end{equation}
\begin{equation}
\chi^{n+\myhalf} = \chi^n + \frac{f}{\overline{\Gamma}_1^{n+1,\star} p_0^{n+1,\star}}\frac{p_{\rm eos}^{n+1,\star}(\rho,h,X_k) - p_0^{n+1,\star}}{\Delta t^n}
\end{equation}
\MarginPar{add volume discrepancy stuff}

@@

\item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
  the updated values $\uadvtwo, \rho_0^{n+1}, (\rho h)_0^{n+1}$, and $p_0^{n+\myhalf}$
  rather than
  $\uadvone, \rho_0^{n+1,\star}, (\rho h)_0^{n+1,\star}$, and $p_0^n$.
  In particular:
\MarginPar{Again, we may not need $\rho_0$ and $(\rho h)_0$.}

@@

\item Update the temperature using the equation of state:
$T^{(2)} =
   T(\rho^{(2)}, h^{(2)}, X_k^{(2)})$ (planar geometry) or
$T^{(2)} =
   T(\rho^{(2)}, p_0^{n+1}, X_k^{(2)})$ (spherical geometry).
\end{enumerate}

@@

First, we compute the time-centered edge velocities, $\Ub^{\nph,\pred}$.
Then, we define
\begin{equation}
\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
\end{equation}
We update the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
equation (\ref{eq:u_update}) as
\begin{eqnarray}
\Ub^{n+1,\dagger}
&=& \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right] \nonumber \\
&& - \dt \left[ - \frac{1}{\rho^\nph} \mathbf{G} \pi^\nmh + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],\nonumber \\
\end{eqnarray}
\MarginPar{I don't think there should be a minus sign in front of the $-\frac{1}{\rho}\mathbf{G} \pi$ term as this has been moved outside the brackets.}
where $\mathbf{G}$ approximates a cell-centered gradient from nodal
data.  Again, the $\dagger$ superscript refers
to the fact that the updated velocity does not satisfy the divergence
constraint.

Finally, we use an approximate nodal projection to define $\Ub^{n+1}$
from $\Ub^{n+1,\dagger},$  such that $\Ub^{n+1}$ approximately
satisfies
\begin{equation}
\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right)
= \beta_0^{n+1} S^{n+1},
\end{equation}
\MarginPar{I think we need the 1/gamma1bar part too}
where
\begin{equation}
\beta_0^{n+1} = \beta \left(\rho_0^{n+1}, p_0^{n+1}, \overline{\Gamma}_1^{n+1}, g^{n+1}\right).
\end{equation}
As part of the projection we also define the new-time perturbational pressure,
$\pi^\nph.$  This projection necessarily differs from the MAC projection used in
{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
on edges and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
and gradient operators.  Details of the approximate projection are given in Paper III.

>>>>

\subsection{Motivation}
We would like to revamp our previous MAESTRO code by moving to the AMReX framework
so that we can take advantage of all of the latest features of the AMReX libraries. 
In addition, we want to modify and simplify the existing numerical scheme 
to both encourage new users and to develop a method that will lend itself more 
easily to incorporating SDC and other high-order methods. 
Irregular base state spacing is considered in order to improve spatial resolution 
that have been seen in the previous MAESTRO code in some test cases.

@@

\item[Step 0.] {\em Initialization}
The initialization step only occurs at the beginning of the simulation.
The initialization step computes initial values for $\pi^{-\myhalf}$, , $\eta_{\rho}^{-\myhalf}$, 
$\psi^{-\myhalf}$, $S^0$, and $S^1$ for use in time stepping routine.  
We pass through {\bf Step 1} -- {\bf Step 12}, but only keep these values.
The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0,
\rho_0^0, p_0^0$, and $\overline{\Gamma}_1^0$ are specified from the problem-dependent
initial conditions.  The initial time step, $\dt^0$, is computed as in
Paper III.

@@

Using the equation below,
%
\begin{equation}
\frac{\partial\Ub}{\partial t} = -\Ub\cdot\nabla\Ub - \frac{1}{\rho}\nabla\pi - \frac{\rho-\rho_0}{\rho}g\eb_r, \label{eq:u_update}
\end{equation}
%
we compute time-centered edge velocities, $\uadvonedag$, using
$\Ub^n$.  The $\dagger$ superscript refers to the
fact that the predicted velocity field does not satisfy the divergence
constraint.  We then construct $\uadvone$ from $\uadvonedag$
using a MAC projection.
We note that $\uadvone$ satisfies the divergence constraint
\begin{equation}
\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \frac{1}{\overline{\Gamma}_1^np_0^n}\psi^{n-\myhalf} + \chi^n \right),
\end{equation}
\begin{equation}
 \beta_0^n = \beta_0 \left(\rho_0^n, p_0^n, \overline{\Gamma}_1^n\right),
\end{equation}
\begin{equation}
\chi^n = \frac{f}{\overline{\Gamma}_1^n p_0^n}\frac{p_{\rm eos}^n(\rho,h,X_k) - p_0^n}{\Delta t^n}
\end{equation}
where $\beta_0$ is computed as described in Appendix C of Paper III.

It should also be noted that we have found it better to approximate 
the temporal changes in base pressure as 
\begin{equation}
\frac{\partial p_0}{\partial t} \approx \psi = \eta_{\rho} g
\end{equation}
instead of using the lagged pressure changes from previous time $t^{n-1}$ 
in the predictor step.

@@

\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$.
\begin{description}
\item For planar geometry,
\begin{eqnarray}
(\rho h)^{(2),\star}
&=& (\rho h)^{(1)} \nonumber \\
&&- \dt^n \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \dt^n \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} \nonumber \\
&& + \dt^n \psi^{n-\myhalf},
\end{eqnarray}

\item For spherical geometry,
\begin{eqnarray}
(\rho h)^{(2),\star}
&=& (\rho h)^{(1)} \nonumber \\
&&- \dt^n \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \dt^n \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) - p_0^{n} \nabla \cdot \uadvone \right \} \nonumber \\
&& + \dt^n \psi^{n-\myhalf},
\end{eqnarray}
\end{description}

@@

Then, for each Cartesian cell where $\rho^{(2),\star} < \rho_\mathrm{cutoff}$,
we recompute enthalpy using
\begin{equation}
(\rho h)^{(2),\star} = \rho^{(2),\star}h\left(\rho^{(2),\star},p_0^{n+1,\star},X_k^{(2),\star}\right).
\end{equation}

@@

The procedure to construct $\uadvtwodag$ is identical to the procedure
for computing $\uadvonedag$ in {\bf Step 3}, but uses
the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
Also, the temporal changes of $p_0$ are now computed directly instead of 
approximated using $\psi$.
We note that $\uadvtwo$ satisfies the divergence constraint
\begin{equation}
\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph} \left(S^{\nph} - \frac{1}{\overline{\Gamma}_1^{n+\myhalf}p_0^{n+1,\star}}\frac{p_0^{n+1,\star}-p_0^n}{\Delta t^n} + \chi^{n+\myhalf}\right),
\end{equation}
\begin{equation}
\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2};
\qquad
 \beta_0^{n+1,\star} = \beta_0 \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \overline{\Gamma}_1^{n+1,\star}\right),
\end{equation}
\begin{equation}
\overline{\Gamma}_1^{\nph} = \frac{ \overline{\Gamma}_1^n +  \overline{\Gamma}_1^{n+1,\star} }{2}.
\end{equation}
\begin{equation}
\chi^{n+\myhalf} = \chi^n + \frac{f}{\overline{\Gamma}_1^{n+1,\star} p_0^{n+1,\star}}\frac{p_{\rm eos}^{n+1,\star}(\rho,h,X_k) - p_0^{n+1,\star}}{\Delta t^n}
\end{equation}

@@

\item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
  the updated values $\uadvtwo, \rho_0^{n+1}, (\rho h)_0^{n+1}$, and $p_0^{n+\myhalf}$
  rather than
  $\uadvone, \rho_0^{n+1,\star}, (\rho h)_0^{n+1,\star}$, and $p_0^n$.
  Also, the temporal changes of $p_0$ are now computed directly instead of 
  approximated using $\psi$.
  In particular:
\MarginPar{Again, we may not need $\rho_0$ and $(\rho h)_0$.}

@@

\item Update the temperature using the equation of state:
$T^{(2)} =
   T(\rho^{(2)}, h^{(2)}, X_k^{(2)})$ (planar geometry) or
$T^{(2)} =
   T(\rho^{(2)}, p_0^{n+1}, X_k^{(2)})$ (spherical geometry).

\item Define a radial cell-centered $\etarho^{\nph}$.

\begin{description}
\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
\begin{equation}
 \etarho^{\nph} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
\end{equation}
\item For spherical geometry, first construct 
$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
\begin{equation}
\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
\end{equation}
Then,
\begin{equation}
\etarho^{\nph,\star} = {\rm {\bf Avg}}\left(\etarho^{{\rm cart},\nph}\right).
\end{equation}
\end{description}

\item Compute $\psi^{\nph}$: 
\begin{equation}
\psi_j^{\nph} = \eta_{\rho,j}^{\nph} g.
\end{equation}

@@

First, we compute the time-centered edge velocities, $\Ub^{\nph,\pred}$.
Then, we define
\begin{equation}
\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
\end{equation}
We update the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
equation (\ref{eq:u_update}) as
\begin{eqnarray}
\Ub^{n+1,\dagger}
&=& \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right] \nonumber \\
&& - \dt \left[ \frac{1}{\rho^\nph} \mathbf{G} \pi^\nmh + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],\nonumber \\
\end{eqnarray}
where $\mathbf{G}$ approximates a cell-centered gradient from nodal
data.  Again, the $\dagger$ superscript refers
to the fact that the updated velocity does not satisfy the divergence
constraint.

Finally, we use an approximate nodal projection to define $\Ub^{n+1}$
from $\Ub^{n+1,\dagger},$  such that $\Ub^{n+1}$ approximately
satisfies
\begin{equation}
\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right)
= \beta_0^{n+1} S^{n+1},
\end{equation}
where
\begin{equation}
\beta_0^{n+1} = \beta \left(\rho_0^{n+1}, p_0^{n+1}, \overline{\Gamma}_1^{n+1}, g^{n+1}\right).
\end{equation}
As part of the projection we also define the new-time perturbational pressure,
$\pi^\nph.$  This projection necessarily differs from the MAC projection used in
{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
on edges and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
and gradient operators.  Details of the approximate projection are given in Paper III.
